A Motion Blur Resilient Fiducial Marker For Quadcopter Imaging

### 2014-03-03 ECCV/paper.tex
added conclusion, corrected typos.

The recent availability of low-cost quadcopters has help to fuel significant efforts in
research focused on Unmanned Aerial Vehicles.  Navigation and planning of
these vehicles is typically performed using onboard inertial sensors and/or
vision based modules that uses visual cues in the real world\cite{Davison:2007}. 
However, to evaluate the effectiveness of navigation methods, artificial
fiducial markers are commonly introduces into the environment to provide
additional information that often serves for ground-truth
measurements(e.g.\cite{Bosnak:2012}\cite{Lim09})

In order to be effective, fiducial markers (or simply called, fiducials) need
to be easily detected in the scene.  There are various existing commonly
used fiducial markers 
\cite{NaimarkF02}\cite{ARToolkit02}\cite{Fiala05}\cite{Pitag13}\cite{runetag11}.
These typically take the form of binary codes arranged into rectangular grids (\cite{ARToolkit02}\cite{Fiala05})
or other geometric features arranged in unique spatial patterns
(\cite{NaimarkF02}\cite{Pitag13}\cite{runetag11}).
Figure \ref{fig:teaser} shows an example of the popular ARTag\cite{Fiala05} as
seen from a quad-copter.  One problem, however, is that low-cost quadcopters
often exhibit very quick and erratic motion.  These fast physical movements
results in motion blur in the quadcopters onboard camera.  Such motion blur has an adverse effect the
reconition of fiducial  markers.  This can be seen in Figure \ref{fig:teaser}(b)
where the ARTag cannot be recongized. This is not too surprising as most
fidicuals are not explicitly designed to handle motion blur.  Compounding this
problem is the common problem of dropped video frames from the quadcopters
wireless communication module.   This means that no only is blurring
introduced, but there may be large discountinues in the patterns position due
to missing video frames that makes tracking challenging.

To address this problem, propose a fiducial design that is robust to motion
blur and drastic drastic change in its position in successive frames.  Our
design is based on concentric circles as shown in Figure \ref{fig:fiducials}. 
The design is based on the observation that motion blur from a quadcopter tends
to be linear in nature.  As such, when our code is blurred, there is no blur in
the direction perpendicular to the direction of motion.   This allows the
signature of the fiducial to remain intact in any direction.  In addition, by
using concentric rings, we can treat them as bits allowing us to encode us
different fiducial identifers.  As far as we are aware, this is the first work
to design a blur resistent marker.   Our experiments show that this design can
significantly outperform existing codes in the face of motion blur.

@@

Intitially there were attempts to design a fiducial embeded with two dimensional
barcode inside a rectangular grid. One of the examples of such fiducial is,
ARToolkit\cite{ARToolkit02}, well known toolkit in Augmented Reality(AR) system. 
Kato et al.\cite{kato-artoolkit} have deomstrated the use of ARToolkit to find
the pose in cideo conferencing system. Many AR applications used ARToolkit
due to its simplicity and efficiency.

Fiala et al. \cite{Fiala05} developed a fiducial termed, ARTag, which is a
bi-tonal system consisting of a square border and an interior region filled
with a 6x6 grid of black or white cells. The improvement in ARTag compared
ARToolkit was detection of coreners instead of detection of lines to find
possible pattern. It proved to be more efficient than \cite{ARToolkit02} in
terms of marker recognition rate as well as the number of different patterns
which can be created. But, drawback of ARTag is, under significant amount of
motion blur, corners can not be detected, resulting in non-recognition of tag.

There were also attempts to use circular patterns instead of rectangular, as
the projective transfomration of circles, i.e. conics, are invariant with
resepctive to point of view of camera\cite{runetag11}. Also, such patterns are
mostly distingushable from background and generally can not be missed.

Concentric rings are reportedly used first by Gatrell et al.\cite{concentric}
for monocular pose estimation as well as object identification in space. Cho et al.
\cite{Cho:2001}\cite{Cho97fastcolor} have used multi color rings instead of
black and white rings\cite{concentric} to increase possible number of fiducials.
These multicolor rings are used in wide area tracking in large scale
applications.

@@

All of these fiducials are not recognizable under occlusion.
Bergamasco et al. \cite{runetag11} have divided fiducial into number of
circular dots, and arranged them in circular fashion, named as RUNE-Tag.
RUNE-tags are reportedly detectable even if 50\% of fiducial area is occluded.
Bergamasco et al. have devloped another fiducial, named as Pi-Tag
\cite{Pitag13}, using cross-ratio concept. As cross-ratio is projective
invariant, the detection of Pi-tag is easier than other tags. Also, compared
to RUNE-tag, it uses remarkably less number of circular dots (12 dots instead of
minimum of 43 dots), still reconisable under occlusion.

@@

Most tracking methods( \cite{Ross:2008} \cite{Wu:2009} \cite{Perez02}
\cite{Mei:2009} ) assume image sequence to be blur free. But in reality, motion
blur is inherent part of most of the videos. Wu et al.\cite{Wu:2011} have
developed BLUr-driven Tracker (BLUT) framework for tracking motion-blurred
targets. BLUT is based on the observation that although motion blurs degrade
the visual features of the target, they, at thde same time, provide useful cues
about the movements to help tracking.

BLUT framework successfully tracks blurred target when there is uniform motion
and the position of tracked object does not change drastically in successive
frames. But in our case, due to frequntly dropping of intermediate frame, this
criterion may not be satisfied.

@@

We begin by first motivating the need for a new blur resistent fiducial by examining
the performance of prior fiducials under motion blur.  After this, we detail
our design as well as the detection algorithm used to find our marker in an image.

@@

We have also analysed the recognition rate along different blur orientations
(from 0 to 90 degrees) at different blur scales. We found that per blur scale,
at certain blur orientations, recognition of ARTag fails. Table shows the range
of blur angles for which we are able to recognise the ARTag at different blur
scales. From Table \ref{tab:artag_blurangle}, it can be seen that, when blur
scale is greater than 30, ARTag is not recognisable with horizontal blur
orientation(blur angle near zero degrees) as well as vertial blur
orientation(blur angle near 90 degrees).

@@

\section{Experimental Validation}
We have implemented our algorithm in C++ using OpenCV library. Our code was run
on a PC with Intel Core i7 processor(@3.4GHz) and 4GB RAM.

@@

\subsection{Comparison}
We will compare our results with standard fiducials such as ARTag. Also, we will
compare our results with Blur driven tracker(BLUT)\cite{Wu:2011}.
\subsubsection{Comparison with ARTag}
First, we have repeated blur simulation experiment on our fiducials too, i.e.,
we have downsized our fiducial to size 150x150 and blurred it along various
orientations with different blur scales. Then, we tried to detect the patterns
using algorithm presented in earlier section. The comparison of recongition
rate is shown in Figure \ref{fig:recognition_rate}.

@@

Later, we recorded the real data using AR Drone quadcopter. In our experimental
setup, we have kept our fiducial along with two AR Tags, to compare the
resilience of blur by each fiducial type. We have used ar\_track\_alvar, ROS
Wrapper for ALVAR libray \cite{ros_alvar}, to detect AR Tags from the stream
captured through quadcopter. The comparison of recognition rate is shown in
Table \ref{tab:recongition_accuracy}. Classification accuracy of all fiducials(ARTag
as well as ours) is almost 100\%.

@@

\subsubsection{Comparison with BLUT}
We have used sample image sequence to test the performance of BLUT\cite{Wu:2011}
and compare it with our result on the same image sequence. From Figure
\ref{fig:BLUT_output}, it can be inferred that BLUT is able to track the fiducial when the position of
fiducial doesnot change too much in successive frames. Also, it can be seen
that, once BLUT looses the track of the fiducial, it is not able to recover from
that. But, our algorithm is able to track the fiducial irrespecive of the change
in the position of the fiducial in the image as shown in Figure
\ref{fig:our_output}.

@@

\textbf{False Negatives} :- We found that sometime our detection algorithm is
failed do recognize our fiducial with binary code ``00'' when there is large
amount of blur. This is the reason behing lower recognition rate of ``00''
coded fiducial compared to that of our other fiducial patterns. When we
investigated the problem, we found that, if there is too much blur, in Gabor
output, the innermost ring is unable to give any response along the direction
perpandicular to blur direction. This problem can be resolved by increasing
radius of innermost ring, reducing effect of blur on the innermost ring.

\textbf{Pose Estimation} :- Currently, we are not finding the pose from the
fiducial marker. But, if we put multiple markers (atleast four) in the same
scene, we may be able to find the pose. The current resolution of onboard camera
will be main hurdle to clear before we can effectively use multiple markers in
each scene.

\textbf{Number of Fiducials} :- Currently, compared to AR Tag fiducial markers,
we are able to generate less number of fiducials. Many applications in
robotics (e.g. qudcopter navigation) may not require large number of fiducial
markers as AR Tag. Most of the time, it is sufficient to have 4-6
different fiducial patterns. Still, we may be able to generate more number of
fiducials by using color background insted of black. E.g., if we just use Red,
Green and Blue as background color, number of fiducials will be increased by
factor of three.

\section{Conclusion and Future Work}

>>>>

The recent availability of low-cost quadcopters has help to fuel significant efforts in
research focused on Unmanned Aerial Vehicles.  Navigation and planning of
these vehicles is typically performed using onboard inertial sensors and/or
vision based modules that uses visual cues in the real world\cite{Davison:2007}. 
However, to evaluate the effectiveness of navigation methods, artificial
fiducial markers are commonly introduces into the environment to provide
additional information that often serves for ground-truth
measurements(e.g.\cite{Bosnak:2012}\cite{Lim09}\cite{Klopschitz:2007})

In order to be effective, fiducial markers (or simply called, fiducials) need
to be easily detected in the scene.  There are various existing commonly
used fiducial markers 
\cite{NaimarkF02}\cite{ARToolkit02}\cite{Fiala05}\cite{Pitag13}\cite{runetag11}.
These typically take the form of binary codes arranged into rectangular grids (\cite{ARToolkit02}\cite{Fiala05})
or other geometric features arranged in unique spatial patterns
(\cite{NaimarkF02}\cite{Pitag13}\cite{runetag11}).
Figure \ref{fig:teaser} shows an example of the popular ARTag\cite{Fiala05} as
seen from a quad-copter.  One problem, however, is that low-cost quadcopters
often exhibit very quick and erratic motion.  These fast physical movements
results in motion blur in the quadcopters onboard camera.  Such motion blur has an adverse effect the
recognition of fiducial  markers.  This can be seen in Figure
\ref{fig:teaser}(b) where the ARTag cannot be recognized. This is not too surprising as most
fiducials are not explicitly designed to handle motion blur.  Compounding this
problem is the common problem of dropped video frames from the quadcopters
wireless communication module.   This means that no only is blurring
introduced, but there may be large discontinuities in the patterns position due
to missing video frames that makes tracking challenging.

To address this problem, propose a fiducial design that is robust to motion
blur and drastic drastic change in its position in successive frames.  Our
design is based on concentric circles as shown in Figure \ref{fig:fiducials}. 
The design is based on the observation that motion blur from a quadcopter tends
to be linear in nature.  As such, when our code is blurred, there is no blur in
the direction perpendicular to the direction of motion.   This allows the
signature of the fiducial to remain intact in any direction.  In addition, by
using concentric rings, we can treat them as bits allowing us to encode us
different fiducial identifiers.  As far as we are aware, this is the first work
to design a blur resistant marker.   Our experiments show that this design can
significantly outperform existing codes in the face of motion blur.

@@

Initially there were attempts to design a fiducial embedded with two
dimensional barcode inside a rectangular grid. One of the examples of such fiducial is,
ARToolkit\cite{ARToolkit02}, well known toolkit in Augmented Reality(AR) system. 
Kato et al.\cite{kato-artoolkit} have demonstrated the use of ARToolkit to find
the pose in video conferencing system. Many AR applications used ARToolkit
due to its simplicity and efficiency.

Fiala et al. \cite{Fiala05} developed a fiducial termed, ARTag, which is a
bi-tonal system consisting of a square border and an interior region filled
with a 6x6 grid of black or white cells. The improvement in ARTag compared to
ARToolkit was detection of corners instead of detection of lines to find
possible pattern. It proved to be more efficient than \cite{ARToolkit02} in
terms of marker recognition rate as well as the number of different patterns
which can be created. But, drawback of ARTag is, under significant amount of
motion blur, corners can not be detected, resulting in non-recognition of tag.

There were also attempts to use circular patterns instead of rectangular, as
the projective transformation of circles, i.e., conics, are invariant with
respective to point of view of camera\cite{runetag11}. Also, such patterns are
mostly distinguishable from background and generally can not be missed.

Concentric rings are reportedly used first by Gatrell et al.\cite{concentric}
for monocular pose estimation as well as object identification in space. Cho et al.
\cite{Cho:2001}\cite{Cho97fastcolor} have used multicolor rings instead of
black and white rings\cite{concentric} to increase possible number of fiducials.
These multicolor rings are used in wide area tracking in large scale
applications.

@@

All of these fiducials are not recognizable under occlusion.
Bergamasco et al. \cite{runetag11} have divided fiducial into number of
circular dots, and arranged them in circular fashion, named as RUNE-Tag.
RUNE-tags are reportedly detectable even if 50\% of fiducial area is occluded.
Bergamasco et al. have developed another fiducial, named as Pi-Tag
\cite{Pitag13}, using cross-ratio concept. As cross-ratio is projective
invariant, the detection of Pi-tag is easier than other tags. Also, compared
to RUNE-tag, it uses remarkably less number of circular dots (12 dots instead of
minimum of 43 dots), still recognisable under occlusion.

@@

Most tracking methods( \cite{Ross:2008} \cite{Wu:2009} \cite{Perez02}
\cite{Mei:2009} ) assume image sequence to be blur free. But in reality, motion
blur is inherent part of most of the videos. Wu et al.\cite{Wu:2011} have
developed BLUr-driven Tracker (BLUT) framework for tracking motion-blurred
targets. BLUT is based on the observation that although motion blurs degrade
the visual features of the target, they at the same time, provide useful cues
about the movements to help tracking.

BLUT framework successfully tracks blurred target when there is uniform motion
and the position of tracked object does not change drastically in successive
frames. But in our case, due to frequently dropping of intermediate frame, this
criterion may not be satisfied.

@@

We begin by first motivating the need for a new blur resistant fiducial by examining
the performance of prior fiducials under motion blur.  After this, we detail
our design as well as the detection algorithm used to find our marker in an image.

@@

We have also analysed the recognition rate along different blur orientations
(from 0 to 90 degrees) at different blur scales. We found that per blur scale,
at certain blur orientations, recognition of ARTag fails. Table shows the range
of blur angles for which we are able to recognise the ARTag at different blur
scales. From Table \ref{tab:artag_blurangle}, it can be seen that, when blur
scale is greater than 30, ARTag is not recognisable with horizontal blur
orientation(blur angle near zero degrees) as well as vertical blur
orientation(blur angle near 90 degrees).

@@

\section{Experimental Validation}
We have implemented\footnote{Code as well as experimental data will be made
online soon} our algorithm in C++ using OpenCV library.
Our code was run on a PC with Intel Core i7 processor(@3.4GHz) and 4GB RAM.

@@

\subsection{Comparison}
We will compare our results with standard fiducials such as ARTag. Also, we will
compare our results with Blur driven tracker(BLUT)\cite{Wu:2011}.
\subsubsection{Comparison with ARTag}
First, we have repeated blur simulation experiment on our fiducials too, i.e.,
we have downsized our fiducial to size 150x150 and blurred it along various
orientations with different blur scales. Then, we tried to detect the patterns
using algorithm presented in earlier section. The comparison of recognition
rate is shown in Figure \ref{fig:recognition_rate}.

@@

Later, we recorded the real data using AR Drone quadcopter. In our experimental
setup, we have kept our fiducial along with two AR Tags, to compare the
resilience of blur by each fiducial type. We have used ar\_track\_alvar, ROS
Wrapper for ALVAR library \cite{ros_alvar}, to detect AR Tags from the stream
captured through quadcopter. The comparison of recognition rate is shown in
Table \ref{tab:recongition_accuracy}. Classification accuracy of all fiducials(ARTag
as well as ours) is almost 100\%.

@@

\subsubsection{Comparison with BLUT}
We have used sample image sequence to test the performance of BLUT\cite{Wu:2011}
and compare it with our result on the same image sequence. From Figure
\ref{fig:BLUT_output}, it can be inferred that BLUT is able to track the fiducial when the position of
fiducial does not change too much in successive frames. Also, it can be seen
that, once BLUT looses the track of the fiducial, it is not able to recover from
that. But, our algorithm is able to track the fiducial irrespective of the
change in the position of the fiducial in the image as shown in Figure
\ref{fig:our_output}.

@@

\textbf{False Negatives} :- We found that sometime our detection algorithm is
failed do recognize our fiducial with binary code ``00'' when there is large
amount of blur. This is the reason behind lower recognition rate of ``00''
coded fiducial compared to that of our other fiducial patterns. When we
investigated the problem, we found that, if there is too much blur, in Gabor
output, the innermost ring is unable to give any response along the direction
perpendicular to blur direction. This problem can be resolved by increasing
radius of innermost ring, reducing effect of blur on the innermost ring.

\textbf{Pose Estimation} :- Currently, we are not finding the pose from the
fiducial marker. But, if we put multiple markers (at least four) in the same
scene, we may be able to find the pose. The current resolution of onboard camera
will be main hurdle to clear before we can effectively use multiple markers in
each scene.

\textbf{Number of Fiducials} :- Currently, compared to AR Tag fiducial markers,
we are able to generate less number of fiducials. Many applications in
robotics (e.g. quadcopter navigation) may not require large number of fiducial
markers as AR Tag. Most of the time, it is sufficient to have 4-6
different fiducial patterns. Still, we may be able to generate more number of
fiducials by using color background instead of black. E.g., if we just use Red,
Green and Blue as background color, number of fiducials will be increased by
factor of three.

\section{Conclusion and Future Work}
Quadcopters are subject to quick and unstable motions that can cause significant
motion blur in the captured images. It severely affects the detection rate of
existing fiducial markers. We proposed the design of a fiducial that is robust
to motion blur. Our design of contrasting concentric rings is based on the the
observation that the direction perpendicular to the motion blur direction will
be unaffected by the blur and therefore still be recognizable. We have shown
through experimental validation that our fiducial will work under significant
amount of motion blur. When compared to popular fiducials, our fiducial can
withstand around twice the amount of blur.  Though, currently we are not able to
find the pose using single pattern, we may be able to overcome this
shortcoming by using multiple markers in single scene. Also, through faster
implementation of Gabor filter, the fiducial detection can be done in realtime.
